Appendix A - Alternative DSDM formulations

### 2019-10-30 manuscript/appendix_a.Rmd
complete IPM code

>>>>

\begin{align}
a
\end{align}

The below code requires some custom functions and a custom greta operation to perform the integration steps and build the kernels efficiently across 3D arrays representing kernels across multiple sites, and to carry out the interpolation of survival rates between adults and seedlings that is implemented in @merowUsingIntegralProjection2014.

- `matrix_dim`: a positive integer giving the dimension of the matrix used to integrate the IPM kernel
- `X_offsize`: a $N$x$K_{\text{offsize}}$ design matrix for the offspring size environmental sub-model
- `X_surv`: a $N$x$K_{\text{surv}}$ design matrix for the survival environmental sub-models
- `X_grow`: a $N$x$K_{\text{grow}}$ design matrix for the growth environmental sub-model
- `X_seedheads`: a $N$x$K_{\text{seedhead}}$ design matrix for the seedhead environmental sub-model
- `beta_env_offsize`: a length $K_{\text{offsize}}$ vector of environmental regression coefficients for offspring size
- `beta_env_surv_seedling`: a length $K_{\text{surv}}$ vector of environmental regression coefficients for logit-seedling-survival
- `beta_env_surv_adult`: a length $K_{\text{surv}}$ vector of environmental regression coefficients for logit-adult-survival
- `beta_env_grow`: a length $K_{\text{grow}}$ vector of environmental regression coefficients for growth
- `beta_env_seedheads`: a length $K_{\text{seedhead}}$ vector of environmental regression coefficients for log-seadhead production
- `beta_size_surv_seedling`: a scalar coefficient for the effect of size on logit-seedling-survival
- `beta_size_surv_adult`: a scalar coefficient for the effect of size on logit-adult-survival
- `beta_size_grow`: a scalar coefficient for the effect of size on growth
- `beta_size_flower`: a scalar coefficient for the effect of size on logit-flowering-probability
- `alpha_germ`: a scalar parameter for logit-germination-probability
- `alpha_seeds`: a scalar parameter for log-seed-production
- `alpha_flower`: a scalar intercept parameter for on logit-flowering-probability
- `offsize_sd`: a positive scalar parameter for the standard deviation of the offspring size distribution
- `grow_sd`: a positive scalar parameter for the standard deviation of the growth distribution

```{r ipms}
# normalise a matrix to sum to one rowwise *in exponential space*. Ie. return y
# where: exp(y[i, ]) = exp(x[i, ]) / sum(exp(x[i, ]))
log_normalise <- function (x) {
  op <- greta::.internals$nodes$constructors$op
  op("log_normalise", x, tf_operation = "tf_log_normalise")
}

# tensorflow code for log normalisation
tf_log_normalise <- function (x) {
  sums <- tensorflow::tf$reduce_logsumexp(x, axis = 2L, keepdims = TRUE)
  x - sums
}

# multiply matrix B by each slice of the 3D array A in 3rd dimension
sweep_prod_3d <- function(A, B) {
  n <- dim(A)[1]
  m <- dim(A)[2]
  Al <- aperm(A, c(2, 3, 1))
  dim(Al) <- c(m, n * m)
  Bl <- t(B)
  dim(Bl) <- n * m
  C <- sweep(Al, 2, Bl, "*")
  dim(C) <- c(m, m, n)
  aperm(C, c(3, 1, 2))
}

# create 3D kernel array for Gaussian transitions (working with the log density
# for computational stability)
distribution_kernel <- function (mean_mat, sd, sizes, bin_width) {
  var <- sd ^ 2
  log_constant <- -0.5 * log(2 * pi) - log(sd)
  diff <- kronecker(sizes, mean_mat, FUN = "-")
  log_density <- log(bin_width) + log_constant - (diff ^ 2) / (2 * var)
  log_density <- log_normalise(log_density)
  dim(log_density) <- c(dim(mean_mat), ncol(mean_mat))
  log_density <- aperm(log_density, c(1, 3, 2))
  exp(log_density)
}

interpolate_survival <- function (adult_surv, seedling_surv, sizes) {
  range <- 0.728 - 0.182
  weights <- 0.728 / range  - (sizes / range)
  weights[sizes > 0.728] <- 0
  weights[sizes <= 0.182] <- 1
  weighted_seedling_surv <- sweep(seedling_surv, 2, weights, FUN = "*")
  weighted_adult_surv <- sweep(adult_surv, 2, 1 - weights, FUN = "*")
  weighted_seedling_surv + weighted_adult_surv
}

min_size <- 0
max_size <- 5
bounds <- min_size + (0:matrix_dim) * (max_size - min_size) / matrix_dim
sizes <- 0.5 * (bounds[-(matrix_dim + 1)] + bounds[-1])
bin_width <- sizes[2] - sizes[1]

# environmental components of sub models
offsize <- X_offsize %*% beta_env_offsize
grow <- X_grow %*% beta_env_grow
env_surv_seedling <- X_surv %*% beta_env_surv_seedling
env_surv_adult <- X_surv %*% beta_env_surv_adult
env_seedheads <- X_seedheads %*% beta_env_seedheads

# size components of submodels
flower <- ilogit(alpha_flower + sizes * beta_size_flower)
size_surv_seedling <- sizes * beta_size_surv_seedling
size_surv_adult <- sizes * beta_size_surv_adult
size_seedheads <- sizes * beta_size_seedheads

# vital rates at all sites
seedheads <- exp(kronecker(size_seedheads, env_seedheads))
surv_seedling <- ilogit(kronecker(size_surv_seedling, env_surv_seedling))
surv_adult <- ilogit(kronecker(size_surv_adult, env_surv_adult))
germination <- ilogit(alpha_germ)
seeds <- exp(alpha_seeds)

# kernelise growth * survival and recruitment * offspring size, and combine
survival <- interpolate_survival(surv_adult, surv_seedling, sizes)
next_size_density <- distribution_kernel(sizes + grow, grow_sd, sizes, bin_width)
P <- sweep_prod_3d(next_size_density, survival)
recruits <- flower * seedheads * seeds * germ
offsize_density <- distribution_kernel(offsize, offsize_sd, sizes, bin_width)
F <- sweep_prod_3d(offsize_density, recruits)
A <- P + F
```
